---
title: "rawDiag - Diagnostic Plots for Mass Spectrometry Data"
subtitle: "Howto Use Diagnostic Plots for Orbitrap Instruments"
author: "Christian Trachsel, Christian Panse"
date: "`r Sys.Date()`"
output: tint::tintPdf
bibliography: rawDiag.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE, warn = -1)
```

# Intro
[@TrachselJPR]

# Installation


The R package can be installed via the following command line.

```{r eval=FALSE}
devtools::install_github("protViz/rawDiag")
```

```{r}
library(rawDiag)
```

# Read Data 



## Thermo Fisher Scientific's RawFileReader

Most promissing for Orbitrap instrument is to use
the function read.raw which uses the 
RawFileReader libraries^[See [RawFileReader](http://planetorbitrap.com/rawfilereader#.WuM718iLmL4) New RawFileReader from Thermo Fisher Scientific.]
```{r eval=FALSE}
MS <- read.raw(file = "20180306_39_autoQC01.raw"), mono = TRUE)
```

## Use mzR

The code snippet below demonstrate how open file standards through using the
mzR^[See [mzR](http://bioconductor.org/packages/mzR/) parser for netCDF, mzXML, mzData and mzML and mzIdentML files.]
bioconductor package can be used.
```{r}
library(mzR)
rawDiag:::as.rawDiag.mzR
```

# Usage

In the following we display some of our demonstration data shipped with the
package. The following code will make the WU163230 data available.
```{r load}
data(WU163230.RData)
```

The package handles data.frame object. having a defined set column names
```{r}
class(WU163230)
rawDiag:::.getDesiredColNames()
```

Helper-functions test if the data.frame have the right column names.
```{r}
is.rawDiag(WU163230)
```

```{r recursorHeatmap, fig.margin = TRUE, fig.cap = "A precursor heatmap of WU163230", fig.width=3.5, fig.height=2.5, cache=TRUE}
PlotPrecursorHeatmap(WU163230)
```

```{r}
knitr::kable(summary.rawDiag(WU163230))
```

```{r cycletime, fig.margin = TRUE, fig.cap = "A cycle time plot of WU163230.", fig.width=3.5, fig.height=7, cache=TRUE}
library(ggplot2)

gp <- PlotCycleTime(WU163230) + 
  scale_y_continuous(breaks = scales::pretty_breaks(3))
```

print object class

```{r}
  class(gp)
```


```{r cycletime2, fig.margin = TRUE, fig.cap = "A cycle time of one raw file.", fig.width=3.5, fig.height=3.5, cache=TRUE}
library(tidyverse)
gp <- WU163230 %>% 
  filter(filename == "20180125_23_500ng_HeLa_15Hz_DDA.raw") %>% 
  PlotScanTime() 

print(gp + 
        facet_grid(filename + MassAnalyzer + MSOrder ~ .))
```
  
# Plot Functions Overview
t.b.d.

# Session Information
An overview of the package versions used to produce this document are shown below.

```{r}
sessionInfo()
```

