---
title: "XIC using rawDiag"
author: "Christian Panse & Christian Trachsel"
date: "6/29/2018"
output:
  pdf_document: default
  html_document: default
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Requirements

If not already installed please install the packages:
- rawDiag download and install from http://fgcz-ms.uzh.ch/~cpanse/rawDiag_0.0.6.tar.gz
- protViz from https://CRAN.R-project.org/package=protViz
- get the example data from  http://fgcz-ms.uzh.ch/~cpanse/20180220_14_autoQC01.raw

```{r include = FALSE}
library(rawDiag)
library(protViz)
```

Define raw file we work on and we calculate the doubly charged mass of three iRT peptides.

```{r}
#addapt to location where the example data file is located
rawfile <- "/Users/cp/Downloads/20180220_14_autoQC01.raw"  
pepSeq <- c("AGGSSEPVTGLADK", "GDLDAASYYAPVR", "FLLQFGAQGSPLFK")
mass2Hplus <- (parentIonMass(pepSeq) + 1.008) / 2 
```

Extract the XIC of the three defined iRT peptides from the raw file.

```{r fig.height=6}
X <- readXICs(rawfile = rawfile, masses = mass2Hplus)

op <- par(mfrow = c(3, 1), mar=c(4,4,4,1))
rv <- lapply(1:3, function(i){plot(X[[i]], main = pepSeq[i])})
```



load the meta data of the specified raw file
```{r}
S <- read.raw(rawfile)
```

Now we check in which scan the delta of the isolated precursor mass and the calculated precursor mass is smaller as 0.1 Da and return a plot for visual inspection. 

```{r}
op <- par(mfrow = c(3, 1))
rv <- lapply(mass2Hplus, function(m){
  plot(S$StartTime, S$PrecursorMass -  m, pch='.',
       ylim=c(-0.25,0.25), main = paste("[m+2H]2+ =", m));
  abline(h=0, col="#77000088");
})
```


uff...so many scans...locks like we have PRM data. How do we verify the peak produced in the XIC plot with the peptide ID?

let's see what else we have.
  
  
we extract the all scan numbers where our precursor filter is true
```{r, include=TRUE}
idx <- lapply(mass2Hplus, function(m){
  which(abs(S$PrecursorMass - m) < 0.1)
})

# just to be generic correct
scanNumbers <- lapply(idx, function(x){S$scanNumber[x]})
```

we define a set of fragment ions for labeling our spectra (from package protviz)
```{r}

HCDIons <- function (b, y) 
{
    Hydrogen <- 1.007825
    Oxygen <- 15.994915
    Nitrogen <- 14.003074
    # c <- b + (Nitrogen + (3 * Hydrogen))
    # z <- y - (Nitrogen + (3 * Hydrogen))
    # return(cbind(b, y,c ,z))
    return(cbind(b, y))
}

```

Now we extract from the raw file the peak list of all scans which match our prcursor filter, pass these scans to peakplot for labeling the b- and y-Ions and use a simple score (largest number of assigned ions) to select the best matching spectrum for each peptide. The score distribution is ploted for visual inspection and the best matching scan is highligted by a red line.

```{r fig.height=6}
op <- par(mfrow = c(3, 1), mar = c(5,4,4,1))
bestMatchingMS2Scan <- sapply(1:3, function(i){
  PL <- readScans(rawfile, scans = scanNumbers[[i]])
  
  pp <- lapply(PL, function(x){psm(pepSeq[i], x, FUN = HCDIons, plot = FALSE)})
  
  # score <- sapply(pp, function(x){sum(abs(x$mZ.Da.error) < 0.1)}) #find best scoring spectra
  score <- sapply(1:length(pp),
     function(j){
       sum(PL[[j]]$intensity[abs(pp[[j]]$mZ.Da.error) < 0.1])}) #find best scoring spectra
  
  plot(1+score, main = paste(pepSeq[i], max(score, na.rm = TRUE)), log='y')
  bestFirstMatch <- which(max(score, na.rm = TRUE) == score)[1]
  abline(v = bestFirstMatch, col = "red")
  # peakplot(pepSeq[i], PL[[bestFirstMatch]])
  scanNumbers[[i]][bestFirstMatch]
})
```


To check the actual fragment spectra, we plot for each peptide the best matching spectrum and label the corresponding b- and y-Ions.
```{r fig.height=6}
op <- par(mfrow = c(3, 1))

PL <- readScans(rawfile, scans = bestMatchingMS2Scan)

pp <- lapply(1:3, function(i){peakplot(pepSeq[i], PL[[i]], FUN = HCDIons)})

```

Now we can plot the XIC again, but zoom with a range of +/- 5 min arround the best matching scan and we are sure we look at the proper peak.

```{r}
op <- par(mfrow = c(3, 1))

lapply(1:3, function(i){
  t <- S$StartTime[bestMatchingMS2Scan[i]]; 
  plot(X[[i]], xlim = c(t-0.25, t+ 0.25)); 
  abline(v=t, col='red')})
```

```{r}
library(protViz)
library(rawDiag)

foo <- function(rawfile, pepSeq){
  mass2Hplus <- (parentIonMass(pepSeq) + 1.008) / 2 
  X <- readXICs(rawfile = rawfile, masses = mass2Hplus)
  S <- read.raw(rawfile)
  
  idx <- lapply(mass2Hplus, function(m){
    which(abs(S$PrecursorMass - m) < 0.1)
  })
  
  
  scanNumbers <- lapply(idx, function(x){S$scanNumber[x]})
  
  bestMatchingMS2Scan <- sapply(1:length(pepSeq), function(i){
    peakList <- readScans(rawfile, scans = scanNumbers[[i]])
    
    peptideSpecMatch <- lapply(peakList, function(x){psm(pepSeq[i], x,
        FUN = HCDIons, plot = FALSE)})
    
    
    score <- sapply(1:length(peptideSpecMatch), function(j){sum(peakList[[j]]$intensity[abs(peptideSpecMatch[[j]]$mZ.Da.error) < 0.1])}) 
    
    bestFirstMatch <- which(max(score, na.rm = TRUE) == score)[1]
    scanNumbers[[i]][bestFirstMatch]
  })
  
  PL <- readScans(rawfile, scans = bestMatchingMS2Scan)
  
  pp <- lapply(1:length(pepSeq), function(j){
    op<-par(mfrow = c(2,1), mar = c(5,4,4,1))
    peakplot(pepSeq[j], PL[[j]], FUN = HCDIons)
    
    t <- S$StartTime[bestMatchingMS2Scan[j]]; 
    plot(X[[j]], xlim = c(t - 0.25, t + 0.25)); 
    abline(v=t, col='red')
    
    x <- X[j]$times - 0.25
    
  })
  
}
```

```{r}

foo(rawfile, c('GAGSSEPVTGLDAK', 'VEATFGVDESNAK', 'TPVISGGPYEYR',
               'TPVITGAPYEYR', 'DGLDAASYYAPVR', 'ADVTPADFSEWSK',
               'GTFIIDPGGVIR', 'LFLQFGAQGSPFLK', 'AGGSSEPVTGLADK',
               'VEATFGVDESANK', 'TPVISGGPYYER', 'TPVITGAPYYER',
               'GDLDAASYYAPVR', 'DAVTPADFSEWSK', 'TGFIIDPGGVIR',
               'FLLQFGAQGSPLFK'))

```
